---
title: "Insert heading" 
author: "Team REGAN"
fontsize: 12pt
geometry: margin=1in
urlcolor: black
output: pdf_document
header-includes:
- \usepackage{setspace}
- \onehalfspacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, highlight=FALSE, comment=NA, warnings=FALSE,message=FALSE)

```

```{r, echo=FALSE, warning=FALSE,results=FALSE}

library(readxl)  
library(dplyr)   
library(ggplot2)   
library(janitor) 
library(scales)
library(tidyr)
library(reshape)
library(tibble)

file_path <- "D:/OneDrive - University of Virginia/Will M/Undergraduate Research Courses-Grants/Summer 2025/Introduction to Regression Analysis (STAT 3220)/Final Project/3220-Final-Project/2025 County Health Rankings Virginia Data - v3.xlsx"

excel_sheets(file_path)

addl_df   <- read_excel(file_path, sheet = "Additional Measure Data") %>% clean_names()
measure_df <- read_excel(file_path, sheet = "Select Measure Data")    %>% clean_names()

full_df <- addl_df %>% 
  inner_join(
      measure_df %>% 
        select(fips,
               percent_with_access_to_exercise_opportunities,
               everything()),
      by = "fips"
  ) %>% 
  mutate(
      exercise_tertile = ntile(percent_with_access_to_exercise_opportunities, 3) %>% 
                         factor(levels = 1:3, labels = c("Low","Mid","High"))
  ) %>% 
  mutate(
      primary_care_physicians_ratio =
           as.numeric(sub(":1$", "", primary_care_physicians_ratio))
  ) %>% 
  mutate(rural_band = cut(percent_rural,
                          breaks = c(-Inf, 25, 50, 75, 100),
                          labels = c("0–25 %", "26–50 %", "51–75 %", "76–100 %"),
                          right = TRUE, ordered_result = TRUE)
  )
```


\begin{center}
\includegraphics[width=\textwidth]%
  {D:/OneDrive - University of Virginia/Will M/Undergraduate Research Courses-Grants/Summer 2025/Introduction to Regression Analysis (STAT 3220)/Final Project/3220-Final-Project/TitlePageImage.png}
  
\end{center}

\newpage 

## Introduction

## Data Summary

### Data Sources

### Exploratory Data Analysis

```{r, echo=FALSE}
## Use this code  to include the relevant plots from EDA
## This should NOT be a dump of all of your plots and outputs.
## Refer to the Rmarkdown guide for sizing information on your plots
```

### EDA Summary

## Methods and Analysis

## Results

## Conclusions

\newpage

## Appendix A: Data Dictionary

|Variable Name|Abbreviated Name|Description|
|:----:|:----:|:----:|
|  |  |   |

\newpage
## Appendix B: Data Rows
```{r, echo=FALSE}
# Use head() function to output first several rows

```

\newpage
## Appendix C: Final Model Output and Plots
```{r, echo=FALSE, warning=FALSE,results=FALSE}
## This code chunk can be used to execute any functions or summaries that will not be included in your output
```

```{r, echo=FALSE}
## Use this section to output the summary of your final model
## Include the relevant plots from assumptions and your additional techniques
## This should NOT be a dump of all of your plots and outputs.
## Refer to the Rmarkdown guide for sizing information on your plots
```

\newpage
## Appendix D: References

### Background

### Data Sources

### Additional Help
